# Bioassays in fruits

## Preventive treatments

Disease severity analysis (spots per fruit) 

```{r}
source(here::here("setup.R"))
library(lme4)
library(glmmTMB)
library(DHARMa)

theme_set(theme_bw(base_size=12))
```


```{r}
dat <- rio::import("data/bioassay_data.csv") %>%
  mutate_at(vars(fungicide, repetition, day, replicate, fruit), as.factor) %>% 
  mutate(fungicide=fct_relevel(fungicide, "control"))
```


```{r}
dat %>% str
```

Data scheme

```{r}
ftable(xtabs(~ fungicide + day + repetition + replicate, dat))
```

### Disease severity 

```{r}
dat %>% 
  ggplot() + 
  aes(x=day, y=spots_per_fruit) + 
  geom_boxplot(width=.5) + 
  geom_jitter(width=.2, col=2, alpha=.5) + 
  labs(x="Treatment", y="Severity (spots per fruit)") + 
  facet_grid(repetition~fungicide)
```

```{r}
dat %>% 
  count(spots_per_fruit) %>% 
  ggplot() + 
  aes(x=spots_per_fruit, y=n) + 
  geom_col()
```

Model fitting

```{r}
# https://glmmtmb.github.io/glmmTMB/articles/glmmTMB.pdf
fit_zipoisson_prev <- glmmTMB(spots_per_fruit~fungicide*day +
                           (1|repetition/replicate),
               ziformula = ~1, family = poisson,
             dat)
# performance::check_overdispersion(fit_zipoisson)
# simulateResiduals(fit_zipoisson) %>% testResiduals()
```

```{r}
fit_zinbinom_prev <- update(fit_zipoisson_prev,family=nbinom2)
fit_zinbinom1_prev <- update(fit_zipoisson_prev,family=nbinom1)
fit_twediee_prev <- update(fit_zipoisson_prev,family=tweedie)
AIC(fit_zipoisson_prev,fit_zinbinom_prev,fit_zinbinom1_prev, fit_twediee_prev)
#               df      AIC
# fit_zipoisson 18 3895.200
# fit_zinbinom  19 3489.228
# fit_zinbinom1 19 3418.201
# fit_twediee   20 3405.606
```

Goodness of fit

```{r}
simulateResiduals(fit_twediee_prev) %>% testResiduals()
```

```{r}
car::Anova(fit_twediee_prev)
```

Means comparison test

```{r}
emm_prev <- emmeans(fit_twediee_prev, ~ fungicide|day, type="response") 
res_prev <- cld(emm_prev, alpha=0.05, Letters=letters,  type="response")
res_prev %>% 
   mutate(`%Control`=abs((response/filter(.,fungicide=="Control")%>% pull(response)-1)*100)) 

res %>% 
  ggplot()+
  aes(x=day, y =response)+
  geom_pointrange(aes(ymin=asymp.LCL , ymax=asymp.UCL))+
  geom_col(alpha=.2, width=.2)+
  facet_wrap("fungicide")
```

### Disease incidence

```{r}
dat_inc <- dat %>% 
  filter(!fungicide=="control") %>% 
  group_by(fungicide, repetition, day, replicate, fruit) %>% 
  summarise(diseased = spots_per_fruit >0) %>% 
  group_by(fungicide, repetition, day, replicate) %>% 
  summarise(n= sum(!is.na(diseased)),
            inc_prop=mean(diseased)) %>% 
  data.frame
```

Model 

```{r}
mod_inc <- glmer(inc_prop ~ fungicide*day +
                   (1|repetition/replicate),
                 weights=n,family=binomial, 
                 data=dat_inc)

simulateResiduals(mod_inc) %>% testResiduals()
```

Means comparison test 

```{r}
emm_inc <- emmeans(mod_inc, ~ day|fungicide, type="response") 
res_inc <- cld(emm_inc, alpha=0.05, Letters=letters,  type="response")
res_inc
```

```{r}
res_inc %>% 
  ggplot()+
  aes(x=day, y =prob)+
  geom_pointrange(aes(ymin=asymp.LCL , ymax=asymp.UCL))+
  geom_col(alpha=.2, width=.2)+
  facet_wrap("fungicide")
```

## Curative treatments

```{r}
cura <- rio::import("data/curative.csv") %>% 
   mutate_at(vars(fungicide, repetition, replicate, fruit), as.factor) %>% 
  mutate(fungicide=fct_relevel(fungicide, "Control")) 
```

Data scheme

```{r}
cura %>% str
ftable(xtabs(~ fungicide + repetition + replicate, cura))
```

### Disease severity (spots per fruit) 

```{r}
cura %>% 
  ggplot() + 
  aes(x=fungicide, y=spots_per_fruit) + 
  geom_boxplot(width=.5) + 
  geom_jitter(width=.2, col=2, alpha=.5) + 
  labs(x="Treatment", y="Severity (spots per fruit)") + 
  facet_wrap("repetition")
```

Model fitting

```{r,eval=FALSE}
# https://glmmtmb.github.io/glmmTMB/articles/glmmTMB.pdf
fit_zipoisson <- glmmTMB(spots_per_fruit~fungicide +
                           (1|repetition/replicate),
               ziformula = ~1, family = poisson,
             cura)
# performance::check_overdispersion(fit_zipoisson)
# simulateResiduals(fit_zipoisson) %>% testResiduals()
```

```{r}
fit_zinbinom <- update(fit_zipoisson,family=nbinom2)
fit_zinbinom1 <- update(fit_zipoisson,family=nbinom1)
fit_twediee <- update(fit_zipoisson,family=tweedie)
AIC(fit_zipoisson,fit_zinbinom,fit_zinbinom1, fit_twediee)

#               df      AIC
# fit_zipoisson  8 1270.421
# fit_zinbinom   9 1176.582
# fit_zinbinom1  9 1182.421
# fit_twediee   10 1180.572
```

Goodness of fit

```{r}
simulateResiduals(fit_twediee) %>% testResiduals()
```

```{r}
car::Anova(fit_twediee)
```

Means comparison test

```{r}
emm_cura <- emmeans(fit_twediee, ~ fungicide, type="response") 
res_cura <- cld(emm_cura, alpha=0.05, Letters=letters,  type="response")
res_cura %>% 
    mutate(`%Control`=abs((response/filter(.,fungicide=="Control")%>% pull(response)-1)*100)) 

res_cura %>% 
  ggplot()+
  aes(x=fungicide, y =response)+
  geom_pointrange(aes(ymin=asymp.LCL , ymax=asymp.UCL))+
  geom_col(alpha=.2, width=.2)
```

### Disease incidence

```{r}
cura_inc <- cura %>% 
  # filter(!fungicide=="control") %>% 
  group_by(fungicide, repetition, replicate, fruit) %>% 
  summarise(diseased = spots_per_fruit >0) %>% 
  group_by(fungicide, repetition, replicate) %>% 
  summarise(n= sum(!is.na(diseased)),
            inc_prop=mean(diseased)) %>% 
  data.frame
```

Model 

```{r}
mod_inc_cura <- glmer(inc_prop ~ fungicide +
                   (1|repetition/replicate),
                 weights=n,family=binomial, 
                 data=cura_inc)
```

Goodness of fit

```{r}
simulateResiduals(mod_inc_cura) %>% testResiduals()
```


Means comparison test

```{r}
emm_inc_cura <- emmeans(mod_inc_cura, ~ fungicide, type="response") 
res_inc_cura <- cld(emm_inc_cura, alpha=0.05, Letters=letters,  type="response")
res_inc_cura
```