---
title: "Field experiments"
date: '`r paste("Actualizado en", Sys.Date())`'
---

```{r, warning=FALSE, message=FALSE}
pacman::p_load(
  tidyverse, 
  rio, 
  janitor, 
  lme4, 
  DHARMa, 
  emmeans, 
  multcomp
)
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
theme_set(theme_bw(base_size=12))
```

```{r}
dat <- import("data/field.csv") %>% 
  mutate_at(vars(fungicide, season, tree), as.factor) %>% 
  mutate(fungicide=fct_relevel(fungicide, "Control"))
str(dat)
```

## Data structure

- 2 Experiments conducted in 2020 and 2021
- 3 Fungicide protection treatments: 2 fungicides + non treated control
- 10 trees per treatment

```{r}
ftable(xtabs(~ season + fungicide + tree, dat))
```


```{r}
dat %>% 
  ggplot() + 
  aes(x=fungicide, y=diseased/leaves) + 
  # geom_boxplot(width=.5) + 
  # geom_text(aes(label=tree))+
  geom_jitter(width=.2, col=2, alpha=.5) + 
  labs(x="Treatment", y="Disease incidence") + 
  facet_wrap("season")
```

## Model fitting

We fitted a generalized linear model including fungicide treatment and season (repetition) and their interaction as fixed effects. 

```{r}
mod <- glm(diseased/leaves ~ fungicide * season,
             weights=leaves, family=binomial, dat)
```

Goodness of fit

```{r fig.show='hide'}
testDispersion(mod)
```

```{r}
car::Anova(mod)
```

## Treatment means comparison per season (individual experiments)

```{r}
emm <- emmeans(mod, ~ fungicide|season, type="response") 
res <- cld(emm, alpha=0.05, Letters=letters,  type="response") %>% 
   mutate(`%Control`=abs((prob/dplyr::filter(.,fungicide=="Control")%>% pull(prob)-1)*100)) %>% 
  tibble

res %>% 
  rename(Season="season") %>% 
  ggplot()+
  aes(x=fungicide, y =prob)+
  geom_pointrange(aes(ymin=asymp.LCL , ymax=asymp.UCL))+
  facet_wrap("Season", labeller = label_both) + geom_text(aes(label=str_squish(.group)), angle=90, vjust=-1, hjust=.5, ) + 
  geom_jitter(data=dat%>%rename(Season="season") , aes(y=diseased/leaves), alpha=.2, width=.1) + 
  labs(y= "Brown spot incidence (proportion)", x="")
```
