---
title: "Fungicide efficay In vitro"
date: '`r paste("Actualizado en", Sys.Date())`'
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE, 
                      fig.height = 6, fig.width = 7)
pacman::p_load(
  tidyverse, 
  medrc # EC50 estimations and comparisons
  )
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
f <- function(x, y, z){
  case_when(
    z < pmin(x, y, na.rm = TRUE) ~ "above",
    z > pmax(x, y, na.rm = TRUE) ~ "below",
    between(z, pmin(x, y, na.rm = TRUE), pmax(x, y, na.rm = TRUE)) ~ "within"
  )
}
```

## Mycelial growth 

```{r}
raw <- rio::import("data/mycelial_growth.csv", dec=",")
dat <- raw %>% 
  mutate_at(vars(dose, colony_diameter), as.numeric) %>% 
  mutate_at(vars(fungicide, strain, experiment, plate), as.factor) %>% 
  # mutate(dose_decimal = format(dose, scientific = FALSE) %>% str_squish()) %>% 
  # mutate(dose = as.numeric(dose_decimal)) %>% 
  mutate(curve_id = interaction(fungicide:strain:experiment)) 
```

Data scheme

```{r}
dat %>% str
ftable(xtabs(~ fungicide + strain + experiment + dose, dat))
```


```{r, eval=FALSE}
# per fungicide / strain
dat %>%  
  nest(data = c(dose, colony_diameter)) %>% 
  mutate(mod = map(data, ~broom::tidy(drm(colony_diameter~dose, fct = LL.3(), data = .))))%>%
  unnest(c(mod))%>% 
  tibble()
```


```{r, eval=FALSE}
# per fungicide
dat %>%  
  nest(data = c(-fungicide)) %>% 
  mutate(mod = map(data, ~broom::tidy(drm(colony_diameter~dose, fct = LL.3(), data = .))))%>%
  unnest(c(mod)) %>% 
  tibble()
```

Meta-analysis approach

```{r}
mod_mg <- metadrm(colony_diameter ~ dose, 
               data=dat,
               fct=LL.3(),
               ind=curve_id,
               cid2=fungicide,
               struct="UN")
summary(mod_mg)
```

Estimates 

```{r}
ec50s <- ED(mod_mg, respLev=c(50)) %>% as.data.frame()
# coef_mod_mg <- summary(mod_mg) %>% data.frame %>% 
#   rownames_to_column("param")  %>% 
#   separate(param, c("param", "fungicide"))
# ec50s <- coef_mod_mg %>% filter(param=="e") 
# ec50s
```


Comparing effective dose levels for meta analysis

```{r}
ed_comp <- EDcomp(mod_mg, 
                  percVec=c(50), 
                  percMat=rbind(c(1,1,1,1)), 
                  interval="fieller") %>% 
      data.frame %>% 
  rownames_to_column("comp") %>%
  rowwise() %>% 
  mutate(relative_to_one = f(Lower, Upper, 1))
```

As we compare EC50 ratios between fungicides, if the confidence interval does not contain 1, fungicides differ among them: 

```{r}
ed_comp %>% 
  ggplot()+ 
  aes(x=comp, y=Estimate) + 
  geom_pointrange(aes(ymin = Lower, ymax = Upper, col=relative_to_one))+     
  geom_hline(yintercept = 1, linetype=2)+
  # scale_y_log10()+
  coord_flip() + 
  labs(col="Relative to 1")
```

> Bellis < Merpan  < Timorex < Ziram

```{r}
dat %>% 
  ggplot()+
  aes(x=dose+0.0001, y=colony_diameter, col=fungicide) +
  geom_point(size=1) + 
  geom_smooth(method = drm, 
              method.args = list(fct = L.3()), se = F) +
    geom_smooth(aes(group=curve_id), size=.3, linetype=2,  
              method = drm, 
              method.args = list(fct = L.3()), se = F) +
  labs(title= "", x = "Dose (ppm)",  y = "Colony diameter (cm)") + 
  scale_x_log10() 
```

## Spore germination

```{r}
germi_raw <- raw <- rio::import("data/germination.csv", dec=",")
germi_dat <- germi_raw %>% 
  mutate_at(vars(dose), as.numeric) %>% 
  mutate_at(vars(fungicide, strain, experiment, plate), as.factor) %>% 
  mutate(curve_id = interaction(fungicide:strain:experiment)) %>% 
  mutate(dose_ = dose+0.0001)
```

Data scheme 

```{r}
germi_dat %>% str
ftable(xtabs(complete.cases(germination_perc)~fungicide+dose, data=germi_dat))
ftable(xtabs(complete.cases(germination_perc)~fungicide+dose+strain , data=germi_dat))
```

Meta-analysis 

```{r}
mod_germ <- metadrm(germination_perc ~ dose_, 
               data=germi_dat,
               fct=LL.3(),
               ind=curve_id,
               cid2=fungicide,
               struct="UN")
summary(mod_germ)
```


```{r}
germ_comp <- EDcomp(mod_germ, 
       percVec=c(50), 
       percMat=rbind(c(1,1,1,1)), 
       interval="fieller")  %>% 
    data.frame %>% 
  rownames_to_column("comp") %>%
  rowwise() %>% 
  mutate(relative_to_one = f(Lower, Upper, 1))
# germ_comp
```


```{r}
coef_mod_germ <- summary(mod_germ) %>% data.frame %>% 
    rownames_to_column("param")  %>% 
  separate(param, c("param", "fungicide"))

# ec50_germ <- coef_mod_germ %>% filter(param=="e") 
# ec50_germ
```


```{r}
germ_comp %>% 
  ggplot()+ 
  aes(x=comp, y=Estimate) + 
  geom_pointrange(aes(ymin = Lower, ymax = Upper, col=relative_to_one))+     
  geom_hline(yintercept = 1, linetype=2)+
  coord_flip() + 
  labs(col="Relative to 1")
```

> Bellis < Ziram < Merpan  <  Timorex

```{r}
germi_dat %>% 
  ggplot()+
  aes(x=dose+0.0001, y=germination_perc, col=fungicide) +
  scale_x_log10() +
  geom_point(size=1) + 
  geom_smooth(method = drm, 
              method.args = list(fct = L.3()), se = F) +  
  geom_smooth(aes(group=curve_id), size=.3, linetype=2,  
              method = drm, 
              method.args = list(fct = L.3()), se = F) +
  labs(title= "", x = "Dose (ppm)",  y = "Germination (%)", col= "Fungicide")  
```
